---
title: 'MGCK registry: Progression report'
output: html_document
---
# To test 
Date: `r format(Sys.time(), "%Y-%m-%d")`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 8)
```
  
This is a [MGCK (MG consortium in Korea) web registry](http://mgck.herokuapp.com) progression report.   
Dataset retrieval: 2016-11  

You can download dataset (.dump, .sql, .csv) and script files (.r, rmd) from [here](https://www.dropbox.com/home/mg/MGCK/WebRegistry_201611). 

<!-- Install and load packages.  -->
```{r, message=FALSE, error=TRUE, echo=FALSE}
if (!require(ggplot2)){
  install.packages("ggplot2")
} else require(ggplot2)
if (!require(dplyr)){
  install.packages("dplyr")
} else library(dplyr)
if (!require(reshape2)){
  install.packages("reshape2")
} else library(reshape2)
if (!require(tidyr)){
  install.packages("tidyr")
} else library(tidyr)
if (!require(pander)){
  install.packages("pander")
} else library(pander)
if (!require(RColorBrewer)){
  install.packages("RColorBrewer")
} else library(RColorBrewer)

library(knitr)
```

<!-- Set working directory -->
```{r, message=FALSE, error=TRUE, echo=FALSE}
setwd("/Users/yoonhohong/GitHub/MGCK")
```
    
## List of contents

1. Demographics
2. Diagnosis
3. Course
4. Remission
5. Crisis
6. Treatment
   + 6.1 Thymectomy
   + 6.2 Pyridostigmine 
   + 6.3 Corticosteroids
   + 6.4 IVIG
   + 6.5 Plasmapheresis
7. Outcome 

```{r, echo=FALSE}
date_course_init = "2014-01-01"
```

## 1. Demographics 

<!-- read demographics file. -->
```{r, echo=FALSE}
dem = read.csv("MGCK_demographics_201611.csv", na.strings = c(""))
tot.no.patients = nrow(dem)
tot.no.hospitals = length(unique(dem$hospital))
dem_tbl = sort(table(dem$hospital), decreasing=T) 
```

* Number of patients enrolled: `r tot.no.patients`
* Number of hospitals: `r tot.no.hospitals`

#### 1) Number of patients across centers 

```{r, echo=FALSE}
pander(dem_tbl, missing="-")
```

#### 2) Age distribution 

```{r, echo=FALSE}
dem$date_birth = as.Date(dem$date_birth, format = "%Y-%m-%d")
dem$age = round(as.numeric(difftime(Sys.time(), dem$date_birth, units = "weeks")/52))
cutpoints = unique(c(min(dem$age), 20, 30, 40, 50, 60, 70, 80, max(dem$age)))
dem$age_group = cut(dem$age, breaks = cutpoints, 
                    include.lowest = TRUE, right = F)
## plot: age distribution
p1 = ggplot(dem, aes(x=age_group, fill=gender))
p1 + geom_bar(position = "dodge") + 
  ggtitle("Age distribution") + 
  xlab("Age") + ylab("Number of patients") + theme_light()
```

## 2. Diagnosis 

<!-- preprocessing dx table -->
```{r, echo=FALSE}
dx = read.csv("MGCK_diagnosis_201611.csv", na.strings = c(""))
## antibodies 
## AChRc 
dx$achrc_ab_dx = ifelse(is.na(dx$achrc_ab_dx), 4, dx$achrc_ab_dx)
dx$achrc_ab_dx = factor(dx$achrc_ab_dx, labels  = 
                          c("negative","borderline","positive", "na")) 
# 1=negative, 2=borderline, 3=positive, 4=na
```

<!-- make a function for pie chart labels -->
```{r, echo=FALSE}
pielbls = function(x){ # x should be a table
  lbl = names(x)
  pct = round(x/sum(x)*100)
  lbl = paste(lbl, pct)
  lbl = paste(lbl, "%", sep="")
  return(lbl)
}
```

#### 1) AChR-Ab 

```{r, echo=FALSE, message=FALSE}
sl_achrc = sort(table(dx$achrc_ab_dx), decreasing = T)
# pichart: AChR-Ab
pie(sl_achrc,labels = pielbls(sl_achrc), 
    col=brewer.pal(length(sl_achrc), "Dark2"),
    main=sprintf("AChR-Ab (n=%d)", sum(sl_achrc)))
```

```{r, echo=FALSE, warning=FALSE}
# histogram: AChR-Ab titer 
p2 = ggplot(dx, aes(x=achrc_ab_titer_dx, y=..count..)) # early vs. late-onset? 
p2 + geom_histogram(fill="midnightblue", col="white", binwidth = 1) +  
  ggtitle(sprintf("AChR-Ab titers (n=%d)", length(which(!is.na(dx$achrc_ab_titer_dx))))) + 
  xlab("AChR-Ab titer (nmol/L)") +
  ylab("Count") + 
  theme_light()
```

#### 2) MuSK-Ab

```{r, echo=FALSE}
dx$musk_ab_dx = ifelse(dx$musk_ab_dx %in% c(4,5,6,NA), NA, dx$musk_ab_dx)
# 1=negative, 2=borderline, 3=positive, 4-6=na 
dx$musk_ab_dx = factor(dx$musk_ab_dx, levels=c(1:3), labels=c("negative", "borderline", "positive"))
```

```{r, echo=FALSE}
sl_musk = table(dx$musk_ab_dx)
pie(sl_musk, labels = pielbls(sl_musk), 
    main = sprintf("MuSK-Ab (n=%d)", sum(sl_musk)), 
    col=brewer.pal(length(sl_musk), "Dark2"))
```

#### 3) Thymus
```{r, echo=FALSE}
# thymus: chest CT
levels(dx$thymus_ct) = list(normal = c(1,"atrophy(involution)","normal"), 
                            hyperplasia = c(2,"hyperplasia"),
                            thymoma = c(3,4,"thymoma", "thymic carcinoma"),
                            others = c(5, "others")
)
# thymus: final 
dx$thymus_dx = ifelse(dx$thymus_dx %in% c(5,6,7), 5, dx$thymus_dx)
dx$thymus_dx = factor(dx$thymus_dx)
levels(dx$thymus_dx) = list(normal = 1,
                            hyperplasia = 2,
                            thymoma = c(3,4),
                            others = 5)
```

```{r, echo=FALSE}
sl_thymusCT = sort(table(dx$thymus_ct), decreasing = T)
sl_thymusDX = sort(table(dx$thymus_dx), decreasing = T)

par(mfrow=c(1,2))
pie(sl_thymusCT, labels = pielbls(sl_thymusCT), 
    main = sprintf("Thymus: CT (n=%d)", sum(sl_thymusCT)),
    col = brewer.pal(length(sl_thymusCT), "Dark2"))
pie(sl_thymusDX, labels = pielbls(sl_thymusDX), 
    main = sprintf("Thymus pathology (n=%d)", sum(sl_thymusDX)),
    col = brewer.pal(length(sl_thymusDX), "Dark2"))
```

```{r, echo=FALSE}
# thymoma staging 
levels(dx$staging_thymoma) = list(I = c("I"), 
                                  IIa = c("IIa", "II_a"),
                                  IIb = c("IIb", "II_b"),
                                  III = c("III"),
                                  IVa = c("IVa", "IV_a"),
                                  IVb = c("IVb", "IV_b"))
```

#### 4) Thymoma: Masaoka staging and WHO classification

```{r, echo=FALSE, warning=FALSE}
# staging 
sl_thymomaST = table(dx$staging_thymoma)
# WHO classification
sl_thymomaWHO = table(dx$who_thymoma)

par(mfrow=c(1,2))
pie(sl_thymomaST, labels = pielbls(sl_thymomaST), 
    col = brewer.pal(length(sl_thymomaST), "Dark2"), 
    main=sprintf("Thymoma staging (n=%d)", sum(sl_thymomaST)))
pie(sl_thymomaWHO, labels=pielbls(sl_thymomaWHO), 
    col=brewer.pal(length(sl_thymomaWHO), "Dark2"), 
    main=sprintf("Thymoma WHO classification (n=%d)", sum(sl_thymomaWHO)))
```

## 3. Course

```{r, echo=FALSE}
# course initial 
ci = read.csv("MGCK_course_initial_201611.csv", na.strings = c(""))
ci$type_onset = factor(ci$type_onset, levels=c(1,2), labels=c("ocular", "generalized"))
ci$type_presentation = factor(ci$type_presentation, levels=c(1,2), labels=c("ocular", "generalized"))
```

#### 1) Type: ocular vs. generalized MG (at onset & presentation)

```{r, echo=FALSE, warning=FALSE}
sl_onset = table(ci$type_onset)
sl_presentation = table(ci$type_presentation)

par(mfrow=c(1,2))
pie(sl_onset, labels=pielbls(sl_onset), col=brewer.pal(length(sl_onset), "Dark2"), 
    main=sprintf("Onset (n=%d)", sum(sl_onset)))
pie(sl_presentation, labels=pielbls(sl_presentation), col=brewer.pal(length(sl_presentation), "Dark2"), 
    main=sprintf("Presentation (n=%d)", sum(sl_presentation)))
```

#### 2) MGFA at entry (first visit) and enrollment (to the registry) 

```{r, echo=FALSE}
levels(ci$mgfa_entry) = list(I = "I",
                             II = c("IIa", "IIb"),
                             III = c("IIIa", "IIIb"),
                             IV = c("IVa", "IVb"),
                             V = "V")
levels(ci$mgfa_firstvisit) = list(CSR = "CSR",
                                  PR = "PR",
                                  MM = "MM",
                                  I = "I",
                                  II = c("IIa", "IIb"),
                                  III = c("IIIa", "IIIb"),
                                  IV = c("IVa", "IVb"),
                                  V = "V")

p = ggplot(ci, aes(x=factor(mgfa_entry), fill=factor(mgfa_firstvisit)))
p + geom_bar(aes(y=(..count..)/sum(..count..)), stat = "count") + 
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percent", fill="MGFA class or PIS at enrollment") + 
  ggtitle(sprintf("MGFA class at entry & enrollment (n=%d)", sum(table(ci$mgfa_entry)))) + 
  xlab("MGFA class at entry") + theme_light() + theme(legend.position = "bottom") +
  coord_flip()
```

#### 3) Time course (onset, presentation, entry, dx, enrollment, secondary generalization)

Summary 
```{r, echo=FALSE}
ci$date_entry = as.Date(ci$date_entry, format="%Y-%m-%d")
ci$date_dx = as.Date(ci$date_dx, format="%Y-%m-%d")
ci$date_firstvisit = as.Date(ci$date_firstvisit, format="%Y-%m-%d")
ci$date_onset = as.Date(ci$date_onset, format="%Y-%m-%d")
ci$date_presentation = as.Date(ci$date_presentation, format="%Y-%m-%d")
ci$date_sec_generalization = as.Date(ci$date_sec_generalization, format="%Y-%m-%d")
  
ci$entry2enrollment = round(as.numeric(difftime(ci$date_firstvisit, ci$date_entry, units="weeks"))/4)
ci$onset2presentation = round(as.numeric(difftime(ci$date_presentation, ci$date_onset, units="weeks"))/4)
ci$presentation2entry = round(as.numeric(difftime(ci$date_entry, ci$date_presentation, units="weeks"))/4)
ci$onset2dx = round(as.numeric(difftime(ci$date_dx, ci$date_onset, units="weeks"))/4)
ci$onset2secgen = round(as.numeric(difftime(ci$date_sec_generalization, ci$date_onset, units="weeks"))/4)

ci_sub = ci[,c("onset2presentation","presentation2entry","entry2enrollment","onset2dx","onset2secgen")]
summary(ci_sub)
```

Patients (pt_serial_no) in whom the time is < 0 (should be *input error*)      

* from onset to presentation: `r  ci[which(ci$onset2presentation <0),]$pt_serial_no`   
* from presentation to entry: `r ci[which(ci$presentation2entry <0),]$pt_serial_no`
* from entry to enrollment: `r ci[which(ci$entry2enrollment <0),]$pt_serial_no`
* from onset to diagnosis: `r ci[which(ci$onset2dx <0),]$pt_serial_no`
* from onset to secondary generalization (in cases of ocular onset): `r ci[which(ci$onset2secgen <0),]$pt_serial_no`

Time distribution

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ci_sub_melted = melt(ci_sub) # convert wide to long format 
ci_sub_melted = ci_sub_melted[!is.na(ci_sub_melted$value),]
ci_sub_melted_sub = ci_sub_melted[ci_sub_melted$value>=0,]

tbl = table(ci_sub_melted_sub$variable)

px = pretty(ci_sub_melted_sub$value)

labels = c(onset2presentation = sprintf("onset2presentation (n=%d)", tbl["onset2presentation"]),
           presentation2entry = sprintf("presentation2entry (n=%d)", tbl["presentation2entry"]),
           entry2enrollment = sprintf("entry2enrollment (n=%d)", tbl["entry2enrollment"]),
           onset2dx = sprintf("onset2dx (n=%d)", tbl["onset2dx"]),
           onset2secgen = sprintf("onset2secgen (n=%d)", tbl["onset2secgen"]))

p =  ggplot(ci_sub_melted_sub, aes(x=value, fill=variable))
p + geom_histogram(col="white") + 
  scale_x_continuous(breaks=px) + 
  facet_wrap(~variable, scales="free_y", nrow=2, labeller = labeller(variable=labels)) + 
  xlab("Time (months)") + theme_light() + theme(legend.position = "none") + 
  theme(strip.text.x = element_text(size=10, colour = "black"))
```

#### 4) MG composite score at enrollment 

Number of patients (second row) across the range of MG composite score at enrollment (first row)
```{r, echo=FALSE}
ci$mgcs_firstvisit = factor(ci$mgcs_firstvisit)
pander(summary(ci$mgcs_firstvisit))
```

#### 5) Change of MG composite score during follow-ups 

```{r, echo=FALSE}
# course fu 
cf = read.csv("MGCK_course_fu_201611.csv", na.strings = c(""))
cf$date_visits = as.Date(cf$date_visits, format="%Y-%m-%d")
cf$pt_serial_no = factor(cf$pt_serial_no)
## change of MG composite score 
cf_mgcs = subset(cf, !is.na(mgcs))
```

MG composite scores are available in `r nrow(cf_mgcs)` visits in `r length(unique(cf_mgcs$pt_serial_no))` patients. 

Number of visits (MG composite score are avaiable) and corresponding number of patients.
```{r, echo=FALSE}
pander(table(table(cf_mgcs$pt_serial_no)))
```

When only the visits after `r date_course_init` were considered, 
```{r, echo=FALSE}
cf_mgcs_sub = subset(cf_mgcs, date_visits >= date_course_init)
```

MG composite scores are available in `r nrow(cf_mgcs_sub)` visits in `r length(unique(cf_mgcs_sub$pt_serial_no))` patients. 

Number of visits (MG composite score are avaiable) and corresponding number of patients.
```{r, echo=FALSE}
pander(table(table(cf_mgcs_sub$pt_serial_no)))
```

Changes of MG composite scores in patients whose MG composite scores are available more than once.  

```{r, echo=FALSE}
mgcs_fu = cf_mgcs_sub[c("pt_serial_no", "date_visits", "mgcs")]

mgcs_fu %>%
  group_by(pt_serial_no) %>%
  arrange(pt_serial_no, date_visits) %>%
  mutate(visit_no = rank(date_visits), visit.total = n()) %>%
  mutate(time_int = round(as.numeric(date_visits - date_visits[1])/30))-> mgcs_fu

temp = mgcs_fu[duplicated(mgcs_fu$pt_serial_no),] # patients whose mgcs are available more than once

fu_mgcs = subset(mgcs_fu, mgcs_fu$pt_serial_no %in% temp$pt_serial_no)
fu_mgcs_no = length(unique(fu_mgcs$pt_serial_no))

p = ggplot(fu_mgcs, aes(x=date_visits, y=mgcs, col=factor(pt_serial_no), group=factor(pt_serial_no)))
p + geom_line() + geom_point() + 
  scale_x_date(date_labels = "%Y-%m-%d") + 
  theme_light() + 
  theme(legend.position = "none") +
  ggtitle(sprintf("Change of MG composite score (n=%d)", fu_mgcs_no)) + 
  xlab("Date of Visit") + 
  ylab("MG composite score")
```
  
```{r, echo=FALSE}
p = ggplot(fu_mgcs, aes(x=time_int, y=mgcs, col=factor(pt_serial_no), group=factor(pt_serial_no)))
p + geom_line() + geom_point() + 
  theme_light() + 
  theme(legend.position = "none") +
  ggtitle(sprintf("Change of MG composite score (n=%d)", fu_mgcs_no)) + 
  xlab("Time from enrollment (month)") + 
  ylab("MG composite score")
```

#### 6) Change of MGFA class 

```{r, echo=FALSE}
levels(cf$mgfa) = list(CSR = "CSR",
                       PR = "PR",
                       MM = "MM",
                       I = "I",
                       II = c("IIa", "IIb"),
                       III = c("IIIa", "IIIb"),
                       IV = c("IVa", "IVb"),
                       V = "V")
cf_mgfa = subset(cf, !is.na(mgfa))
```

MGFA class are available in `r nrow(cf_mgfa)` visits in `r length(unique(cf_mgfa$pt_serial_no))` pateints.   

Number of visits (MGFA class is avaiable) and corresponding number of patients.
```{r, echo=FALSE}
pander(table(table(cf_mgfa$pt_serial_no)))
```

When only the visits after `r date_course_init` were considered, 
```{r, echo=FALSE}
cf_mgfa_sub = subset(cf_mgfa, date_visits >= date_course_init)
```

MGFA class (or post-intervention status) is avaiable in `r nrow(cf_mgfa_sub)` visits in `r length(unique(cf_mgfa_sub$pt_serial_no))` patients. 

Number of visits (MGFA class or PIS is avaiable) and corresponding number of patients. 
```{r, echo=FALSE}
pander(table(table(cf_mgfa_sub$pt_serial_no)))
```

Change of MGFA class  

```{r, echo=FALSE}
mgfa_fu = cf_mgfa_sub[c("pt_serial_no", "date_visits", "mgfa")]

library(dplyr)
mgfa_fu %>%
  group_by(pt_serial_no) %>%
  arrange(pt_serial_no, date_visits) %>%
  mutate(visit_no = rank(date_visits), visit.total = n()) %>%
  mutate(time_int = round(as.numeric(date_visits - date_visits[1])/30))-> mgfa_fu

temp = mgfa_fu[duplicated(mgfa_fu$pt_serial_no),]

fu_mgfa = subset(mgfa_fu, mgfa_fu$pt_serial_no %in% temp$pt_serial_no)
fu_mgfa_no = length(unique(fu_mgfa$pt_serial_no))

p = ggplot(fu_mgfa, aes(x=date_visits, y=mgfa, col=factor(pt_serial_no), group=factor(pt_serial_no)))
p + geom_line() + geom_point() + 
  scale_x_date(date_labels = "%Y-%m-%d") + 
  theme_light() + 
  theme(legend.position = "none") +
  ggtitle(sprintf("Change of MGFA class (n=%d)", fu_mgfa_no)) + 
  xlab("Date of Visit") + 
  ylab("MGFA class")
```

```{r, echo=FALSE}
p = ggplot(fu_mgfa, aes(x=time_int, y=mgfa, col=factor(pt_serial_no), group=factor(pt_serial_no)))
p + geom_line() + geom_point() + 
  theme_light() + 
  theme(legend.position = "none") +
  ggtitle(sprintf("Change of MGFA class (n=%d)", fu_mgfa_no)) + 
  xlab("Time from enrollment (month)") + 
  ylab("MGFA class")
```


## 4. Remission

Number of patients who achieved remission  
Time to remission   
Duration of remission   
Number of relapse  
Time to relapse  
  

```{r, echo=FALSE}
rem = read.csv("MGCK_remission_201611.csv", na.strings = c(""))
# number of patients who achieved remission
# time2remission (dx2remission) 
# duration_remission
# number of patients who achieved remission but relapsed thereafter
# time2relapse (remission2relapse)
```

## 5. crisis

```{r, echo=FALSE}
crisis = read.csv("MGCK_crisis_201611.csv", na.strings = c(""))
# number of patients who got crisis
crisis_sub = subset(crisis, !is.na(date_admission_mg_crisis))
```

Number of crisis: `r nrow(crisis_sub)`  
Number of patients who got at least one crisis: `r length(unique(crisis_sub$pt_serial_no))`  

Number of crisis (first row) and correponding number of patients    
```{r, echo=FALSE}
crisis_tbl = table(table(crisis_sub$pt_serial_no))
pander(crisis_tbl)
```

Time from onset to crisis (excluding cases in which time to crisis is longer than 60 months)  

```{r, echo=FALSE}
# time2crisis
crisis_sub$date_admission_mg_crisis = as.Date(crisis_sub$date_admission_mg_crisis, 
                                              format = "%Y-%m-%d")
temp = merge(crisis_sub, ci, by="pt_serial_no")
temp = subset(temp, !is.na(date_onset))
temp %>%
  select(pt_serial_no, date_onset, date_admission_mg_crisis) %>%
  mutate(time2crisis = round(as.numeric(
    difftime(temp$date_admission_mg_crisis, temp$date_onset, units = "weeks")
    )/4)) -> time2crisis_df
```

```{r, echo=FALSE}
time2crisis_df_sub = subset(time2crisis_df, time2crisis <= 60)
time2crisis_df_sub %>%
  group_by(pt_serial_no) %>%
  arrange(date_admission_mg_crisis) -> time2crisis_df_sub

p = ggplot(time2crisis_df_sub, aes(x=factor(pt_serial_no), y=time2crisis))
p + geom_point() + geom_line() + 
  geom_segment(aes(x=factor(pt_serial_no), y=0, xend=factor(pt_serial_no), yend=time2crisis)) + 
  xlab("patient identifier") + ylab("Time from onset to crisis (months)") + theme_light() + 
  coord_flip()
```

ICU stay (days) d/t MG crisis

```{r, echo=FALSE}
# days_icu_stay
crisis_icu = subset(crisis, !is.na(days_icu_stay_mg_crisis))
hist(crisis_icu$days_icu_stay_mg_crisis, main = sprintf("ICU stay (n=%d)", nrow(crisis_icu)), col="darkblue", 
     xlab="Days") 
```
```{r, echo=F}
summary(crisis_icu$days_icu_stay_mg_crisis)
```

## 6. Treatment 

#### 6.1 Thymectomy 

```{r, echo=FALSE}
thymectomy = read.csv("MGCK_thymectomy_201611.csv", na.strings = c(""))
thymectomy$date_thymectomy = as.Date(thymectomy$date_thymectomy)
levels(thymectomy$thymectomy_type_resection) = 
  list(transcervical_basic = c(1, "transcervical_basic"),
       transcervical_extended = c(2, "transcervical_extended"),
       transsternal_classic = c(3, "transsternal_classic"),
       transsternal_extended = c(4, "transsternal_extended"),
       transcervical_transsternal = c(5, "transcervical_transsternal"),
       videoscopic_classic_VATS = c(6, "videoscopic_classic_VATS"),
       videoscopic_extended_VATET = c(7, "videoscopic_extended_VATET"),
       robot_surgery = c(8, "robot_surgery"),
       others = c(9, "others"))
```

```{r, echo=FALSE}
# number of patients who got thymectomy (excluding thymoma)
thymectomy_merged = merge(dx, thymectomy, by="pt_serial_no") 
thymectomy_nonthymoma = subset(thymectomy_merged, !thymus_dx == "thymoma") 
thymectomy_nonthymoma = subset(thymectomy_nonthymoma, !thymus_ct == "thymoma") 
thymectomy_thymoma = subset(thymectomy_merged, thymus_dx == "thymoma"|thymus_ct == "thymoma") 

thymectomy_nonthymoma_sub = subset(thymectomy_nonthymoma, !is.na(date_thymectomy))
thymectomy_thymoma_sub = subset(thymectomy_thymoma, !is.na(date_thymectomy))

```

Number of patients who got thymectomy for non-thymomatous MG: `r nrow(thymectomy_nonthymoma_sub)`
Number of patients who got thymectomy for non-thymomatous MG: `r nrow(thymectomy_thymoma_sub)`

Time from onset to thymectomy for non-thymomatous MG 

```{r, echo=FALSE}
time2thymectomy = merge(thymectomy_nonthymoma_sub, ci, by="pt_serial_no")
time2thymectomy %>%
  mutate(onset2thymectomy = round(as.numeric(difftime(date_thymectomy, date_onset, units = "weeks"))/4)) -> time2thymectomy

hist(time2thymectomy$onset2thymectomy, 
     main=sprintf("Time from onset to thymectomy (in cases of nonthymoma, n=%d)", nrow(time2thymectomy)),
     xlab = "Time from onset to thymectomy (months)", 
     ylab = "No. of patients", col="darkblue")
```

Thymectomy: type of resection  
For non-thymoma, 
```{r, echo=FALSE}
# thymectomy_type_resection
thymectomy_nonthymoma_tbl = table(thymectomy_nonthymoma_sub$thymectomy_type_resection)
thymectomy_thymoma_tbl = table(thymectomy_thymoma_sub$thymectomy_type_resection)
thymectomy_nonthymoma_tbl
```

For thymoma, 
```{r, echo=FALSE}
thymectomy_thymoma_tbl
```

#### 6.2 Pyridostigmine 

```{r, echo=FALSE}
# pyrido
pyrido = read.csv("MGCK_pyrido_201611.csv", na.strings = c(""))
```

#### 6.3 Corticosteroids 

```{r, echo=FALSE}
# steroid 
steroid = read.csv("MGCK_steroid_201611.csv", na.strings=c(""))
steroid$date_start_steroid = as.Date(steroid$date_start_steroid, format="%Y-%m-%d")
steroid$date_end_steroid = as.Date(steroid$date_end_steroid, format="%Y-%m-%d")
levels(steroid$steroid_route) = list(iv = c("iv", "i.v."),
                                     oral = "oral")
```

```{r, echo=FALSE}
steroid_iv = subset(steroid, steroid_route %in% c("iv", "i.v."))
steroid_oral = subset(steroid, steroid_route == "oral")
```

Number of patients treated with iv methylprednisolone: `r length(unique(steroid_iv$pt_serial_no))`
Number of patients treated with oral prednisolone: `r length(unique(steroid_oral$pt_serial_no))`

Changes of oral prednisolone dose   

```{r, echo=FALSE}
steroid_oral_sub = steroid_oral[c("pt_serial_no", "date_start_steroid", 
                                  "date_end_steroid",
                                  "steroid_dose")]  
steroid_dose = steroid_oral_sub[complete.cases(steroid_oral_sub),]
steroid_dose %>%
  group_by(pt_serial_no) %>%
  arrange(pt_serial_no, date_start_steroid) %>%
  mutate(steroid_tot = n()) -> steroid_dose
# wide2long format
steroid_dose_long = gather(steroid_dose, starend_point, date, date_start_steroid:date_end_steroid, factor_key = T)

steroid_dose_long %>%
  group_by(pt_serial_no) %>%
  arrange(pt_serial_no, date) %>%
  mutate(steroid_no = rank(date)) %>%
  mutate(time_int = as.numeric(date - date[1])) -> steroid_dose_long
```

```{r, echo=FALSE}
p = ggplot(steroid_dose_long, aes(x=date, y=steroid_dose, 
                                  col=factor(pt_serial_no), 
                                  group=factor(pt_serial_no)))
p + geom_line() + geom_point() + 
  scale_x_date(date_labels = "%Y-%m-%d") + 
  theme_light() + 
  theme(legend.position = "none") +
  ggtitle(sprintf("Oral prednsiolone (n=%d)", length(unique(steroid_dose_long$pt_serial_no)))) + 
  xlab("Date of Visit") + 
  ylab("Dose (mg/day)")
```

```{r, echo=FALSE}
p = ggplot(steroid_dose_long, aes(x=time_int, y=steroid_dose, 
                                  col=factor(pt_serial_no), 
                                  group=factor(pt_serial_no)))
p + geom_line() + geom_point() + 
  theme_light() + 
  theme(legend.position = "none") +
  ggtitle(sprintf("Oral prednisolone (n=%d)", length(unique(steroid_dose_long$pt_serial_no)))) + 
  xlab("Time (days)") + 
  ylab("Dose (mg/day)")
```

#### 6.4 Immunosuppressive agents   

```{r, echo=FALSE}
# istx 
istx = read.csv("MGCK_istx_201611.csv", na.strings = c(""))
istx$side_effects_is_tx = factor(istx$side_effects_is_tx)
levels(istx$is_tx_route) = list(iv = c("i.v.", "iv"),
                                oral = "oral") 
istx_use = subset(istx, !is.na(is_tx_drug))
```

Number of patients treated with immunosuppressive agents: `r length(unique(istx_use$pt_serial_no))`    
Number of patients (second row) in whom N (first row) different immunosuppressive agents were used in total  
```{r, echo=FALSE}
istx_mat = as.matrix(table(istx_use$pt_serial_no, istx_use$is_tx_drug))
istx_no = apply(istx_mat, 1, function(x)(length(which(x>0))))
pander(table(istx_no)) # number of patients who received N immunosuppressive agents in total 
```

Use of immunosuppressive agents: proportion  

```{r, echo=FALSE}
istx_df = as.data.frame(table(istx_use$pt_serial_no, istx_use$is_tx_drug))
colnames(istx_df) = c("pt_serial_no", "drug", "Freq")

istx_df %>%
  filter(Freq>0) %>%
  group_by(drug) %>%
  summarize(n = n()) -> istx_df 
# number of patients who got treated each immunosuppressive agent at least once

istx_v = istx_df$n
names(istx_v) = istx_df$drug

pie(istx_v, labels=pielbls(istx_v), col=brewer.pal(length(istx_v), "Dark2"), 
    main=sprintf("Immunosuppressive agents (n=%d)", sum(istx_v)))
```
  
```{r, echo=FALSE}
istx_complete = subset(istx, !is.na(date_start_is_tx))
istx_complete = subset(istx_complete, !is.na(date_end_is_tx))

```

Number of patients in whom the dates of start and end are both specified: `r length(unique(istx_complete$pt_serial_no))`  

Side effects of immunosuppressive agents (0; no, 1; yes)   
```{r, echo=FALSE}
pander(table(istx$side_effects_is_tx))   
```



#### 6.5 IVIG 

```{r, echo=FALSE}
# ivig
ivig = read.csv("MGCK_ivig_201611.csv", na.strings = c(""))
ivig_use = subset(ivig, !is.na(date_start_ivig))
```

IVIG was treated in `r nrow(ivig_use)` occasions in `r  length(unique(ivig_use$pt_serial_no))`  patients. 

Number of IVIG use (first row) and corresponding number of patients.
```{r, echo=FALSE}
pander(table(table(ivig_use$pt_serial_no))) # number of ivig tx; number of patients 
```

Number of occasions in which side effects of IVIG were reported: 
```{r, echo=FALSE}
pander(table(ivig_use$side_effects_ivig))
```

#### 6.6 Plasmapheresis

```{r, echo=FALSE}
pex = read.csv("MGCK_pex_201611.csv", na.strings = c(""))
pex_use = subset(pex, !is.na(date_start_pex))
pex_use$side_effects_pex = factor(pex_use$side_effects_pex)
```

Plasmapheresis was used in `r nrow(pex_use)` occasions in `r length(unique(pex_use$pt_serial_no))` patients. 

Number of plasmapheresis (first row) and corresponding number of patients.
```{r, echo=FALSE}
pander(table(table(pex_use$pt_serial_no)))
```

Side effects of PE (0=no; 1=yes)
```{r}
summary(pex_use$side_effects_pex)
```

## 7. Outcome

Follow-up duration: from first visit (enrollment) to date close  
```{r, echo=FALSE}
outcome = read.csv("MGCK_outcome_201611.csv", na.strings = c(""))
ci_sub = ci[c("pt_serial_no", "date_firstvisit", "mgfa_firstvisit", "mgcs_firstvisit")]
outcome = merge(outcome, ci_sub, by="pt_serial_no")
outcome_sub = subset(outcome, !is.na(date_close))
# fu_duraton 
temp = outcome_sub[c("pt_serial_no", "date_firstvisit", "date_close")]
fu_duration_df = temp[complete.cases(temp),]
fu_duration_df$date_close = as.Date(fu_duration_df$date_close)
fu_duration_df %>%
  mutate(fu_duration = 
           round(as.numeric(difftime(date_close, date_firstvisit, 
                                     units = "weeks"))/4)) -> fu_duration_df
summary(fu_duration_df$fu_duration)
```

When only the patients with 0 <= fu duration <= 60 months were considered,    
```{r}
fu_duration_df_sub = subset(fu_duration_df, fu_duration <= 60 & fu_duration >= 0)

p = ggplot(fu_duration_df_sub, aes(x=fu_duration))
p + geom_histogram(fill="darkblue", binwidth = 1) + 
  ggtitle(sprintf("Follow-up duration (n=%d)", nrow(fu_duration_df_sub))) + 
  theme_light() + 
  xlab("Follow-up duration (months)")
```

```{r}
levels(outcome$mgfa_firstvisit) = list(CSR = "CSR", 
                                       PR = "PR", 
                                       MM = "MM",
                                       I = "I",
                                       II = c("IIa", "IIb"),
                                       III = c("IIIa", "IIIb"),
                                       IV = c("IVa", "IVb"),
                                       V = "V")
levels(outcome$mgfa_lastvisit) = list(CSR = "CSR",
                                  PR = "PR",
                                  MM = "MM",
                                  I = "I",
                                  II = c("IIa", "IIb"),
                                  III = c("IIIa", "IIIb"),
                                  IV = c("IVa", "IVb"),
                                  V = "V")
```

MGFA class at the end of f/u  

```{r}
outcome_mgfa = subset(outcome, !is.na(mgfa_lastvisit))
p = ggplot(outcome_mgfa, aes(x=factor(mgfa_lastvisit), fill=factor(mgfa_firstvisit)))
p + geom_bar(aes(y=(..count..)/sum(..count..)), stat = "count") + 
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percent", fill="MGFA class or PIS at enrollment") + 
  ggtitle(sprintf("MGFA class at the end of fu (n=%d)", sum(table(outcome_mgfa$mgfa_lastvisit)))) + 
  xlab("MGFA class at the end of fu") + theme_light() + theme(legend.position = "bottom") +
  coord_flip()
```

MG composite scores at the end of f/u   

```{r, warning=FALSE}
outcome_mgcs = outcome[complete.cases(outcome[c("mgcs_firstvisit", "mgcs_lastvisit")]),]
temp = outcome_mgcs[c("pt_serial_no", "mgcs_firstvisit", "mgcs_lastvisit")]

temp = gather(temp, visit, mgcs, mgcs_firstvisit:mgcs_lastvisit, factor_key = T)

pd = position_dodge(0.4)
p = ggplot(temp, aes(x=visit, y=jitter(as.numeric(mgcs)), col=factor(pt_serial_no), 
                     group=factor(pt_serial_no)))
p  + geom_line(position = pd, alpha=.5) + 
  geom_point(position = pd, alpha=.5) +
  theme_light() + scale_x_discrete(labels=c("At enrollment", "At the end of f/u")) + 
  theme(legend.position = "none",
        axis.title.x = element_blank()) + 
  ylab("MG composite score") + 
  ggtitle(sprintf("Changes of MG composite score (n=%d)", length(unique(outcome_mgcs$pt_serial_no))))
```

If you have any question or suggestion, contact me at nrhong@gmail.com

                                            -- The End --


  


